---
title: "Untitled"
author: "Laura Menicacci"
date: "2023-02-26"
output: html_document
---

# Libraries & data

```{r setup, include=FALSE}

library(tidyverse)
pct_patent_counts <- read_csv("D:/pct_patent_counts_1.csv")

```

# Introduction 

* Why I chose PCT Office


# Add 0 values for missing years

Here I assume that years that are not present 

```{r}

years = (1977:2020)

#create data frame with 0 rows and 3 columns
df2 <- data.frame(matrix(ncol = 3, nrow = 0))

# provide column names

for (state in unique(pct_patent_counts$Up_reg_label)){
  for (i in (1:length(years))){
    df2 <- rbind(df2, c(state, years[i], 0))
  }
}
colnames(df2) <- c('Up_reg_label','prio_year','inv_share')
for(i in (1:nrow(pct_patent_counts))){
  for (j in (1:nrow(df2))){
    if(pct_patent_counts[i,1] == df2[j,1] && pct_patent_counts[i,2] == df2[j,2]){
      df2[j,3] = pct_patent_counts[i,3]
    }
  }
}


```

# Sistema variabili

```{r}

df2$inv_share <- as.numeric(df2$inv_share)

pat <- df2

pat <- pat %>% 
  rename(region = Up_reg_label) %>% 
  rename(patent_count = inv_share) %>% 
  as.integer(prio_year)

write_csv(pat,"D:\\pct_patent_counts_1.csv")

```

# Summary statistics

```{r}
pct_patent_counts %>% summarise(across(where(is.numeric), .fns = 
                     list(min = min,
                          median = median,
                          mean = mean,
                          stdev = sd,
                          q25 = ~quantile(., 0.25),
                          q75 = ~quantile(., 0.75),
                          max = max))) %>%
  pivot_longer(everything(), names_sep='_', names_to=c('variable', '.value'))

```
# Filter for specific time frame

```{r}
pct_pat <- pct_patent_counts %>% 
  filter(prio_year >= 1999 & prio_year <= 2020)

write_csv(pct_pat,"D:\\pct_patent_counts_1999_2020.csv")
```


# Add GDP control 

Real GDP (millions of chained 2012 dollars)

Unit: Millions of chained 2012 dollars

Source: BEA 
```{r}

gdp <- read_csv("D:/Controls/SAGDP1__ALL_AREAS_1997_2021.csv")

gdp_filt <- gdp %>% 
  select(!c("GeoFIPS", "Region", "TableName", "LineCode", "IndustryClassification")) %>% 
  filter(Description %in% c("Real GDP (millions of chained 2012 dollars)")) %>% # "Current-dollar GDP (millions of current dollars)"
  filter(!GeoName %in% c("United States")) %>% 
  pivot_longer(cols = c("1997", "1998", "1999","2000" ,"2001" ,"2002" ,"2003" ,"2004" ,"2005" ,"2006" ,"2007" ,"2008" ,"2009" ,"2010" ,"2011" , "2012","2013","2014","2015","2016","2017","2018","2019","2020", "2021"), 
               names_to = "year", 
               values_to = "gdp") %>% 
  select(!c("Description", "Unit")) 

gdp_filt$year <- as.numeric(gdp_filt$year)

pat_gdp <- left_join(pct_pat, gdp_filt, by = c("region" = "GeoName", "prio_year" = "year"))

write_csv(pat_gdp,"D:\\pat_gdp.csv")

```

# Population 

Resident population including Armed Forces, 

Unit of Measure: thousand

Sources: Table by the U.S. Energy Information Administration, based on population data from the U.S. Census Bureau

```{r}
library(readxl)
pop <- read_excel("D:\\Controls\\use_oi.xlsx", sheet = 2, skip = 2)

us_codes <- as.data.frame(state.abb)

us_codes$name <- state.name

pat_codes <- left_join(pat_gdp, us_codes, by = c("region" = "name"))

pop_filt <-  pop[,c(1,39:62)]

pop_long <- pop_filt %>% pivot_longer(cols = c("1997", "1998", "1999","2000" ,"2001" ,"2002" ,"2003" ,"2004" ,"2005" ,"2006" ,"2007" ,"2008" ,"2009" ,"2010" ,"2011" , "2012","2013","2014","2015","2016","2017","2018","2019","2020"), 
               names_to = "year", 
               values_to = "pop")

pop_long$year <- as.numeric(pop_long$year)

pat_gdp_pop <- left_join(pat_codes, pop_long, by = c("state.abb" = "State", "prio_year" = "year"))

pat_gdp_pop <- pat_gdp_pop %>% 
  rename("state_name"= "region") %>% 
  rename("state_code" = "state.abb") %>% 
  select(c("state_name", "state_code", "prio_year", "patent_count", "gdp", "pop"))

```


# R&D control

R&D and R&D Plant Obligations by U.S. State

The amount of federal obligations for R&D and R&D plant to U.S. performers by state.
Obligations represent the amounts for orders placed, contracts awarded, services received, and similar transactions during a given period, regardless of when the funds were appropriated.

Deflator: Current Dollars			
Unit of Measure: Thousands of Dollars			
SOURCE: National Center for Science and Engineering Statistics, Survey of Federal Funds for Research and Development.	

```{r}
r_d <- read_csv("D:/Controls/r&d_exp_1999_2019_new.csv")

r_d_long <- r_d %>% 
  pivot_longer(cols = new,  
               names_to = "year", 
               values_to = "rd", values_transform = list(new = as.numeric))

r_d_long$year <- as.numeric(r_d_long$year)

pat_gdp_pop_rd <- left_join(pat_gdp_pop, r_d_long, by = c("state_name" = "Fiscal Year", "prio_year" = "year"))

write_csv(pat_gdp_pop_rd,"D:\\pat_gdp_pop_rd.csv")

```


# Crude oil prices

Numbers aggregated by Petroleum Administration for Defense District (PADD):

PADD 1 (East Coast):
   PADD 1A (New England): Connecticut, Maine, Massachusetts, New Hampshire, Rhode Island, Vermont.
   PADD 1B (Central Atlantic): Delaware, District of Columbia, Maryland, New Jersey, New York, Pennsylvania.
   PADD 1C (Lower Atlantic): Florida, Georgia, North Carolina, South Carolina, Virginia, West Virginia.
PADD 2 (Midwest): Illinois, Indiana, Iowa, Kansas, Kentucky, Michigan, Minnesota, Missouri, Nebraska, North Dakota, Ohio, Oklahoma, South Dakota, Tennessee, Wisconsin.
PADD 3 (Gulf Coast): Alabama, Arkansas, Louisiana, Mississippi, New Mexico, Texas, Federal Offshore Gulf.
PADD 4 (Rocky Mountain): Colorado, Idaho, Montana, Utah, Wyoming.
PADD 5 (West Coast): Alaska (North Slope and Other Mainland), Arizona, California, Hawaii, Nevada, Oregon, Washington, Federal Offshore California.
```{r}

library(readxl)
fuel <- read_excel("D:\\Controls\\fuel_prices\\PET_PRI_DFP1_K_A.xls", sheet = 2, skip = 2)

fuel_filt <- fuel %>% 
  select(c("Date", "U.S. Crude Oil First Purchase Price (Dollars per Barrel)", "East Coast (PADD 1) Crude Oil First Purchase Price (Dollars per Barrel)", "Midwest (PADD 2) Crude Oil First Purchase Price (Dollars per Barrel)", "Gulf Coast (PADD 3) Crude Oil First Purchase Price (Dollars per Barrel)", "Rocky Mountain (PADD 4) Crude Oil First Purchase Price (Dollars per Barrel)", "West Coast (PADD 5) Crude Oil First Purchase Price (Dollars per Barrel)"))

# change date format to year w lubridate

# how to go from padd to countries???


```

