---
title: "Summary statitics + controls"
author: "Laura Menicacci"
date: "2023-02-26"
output:
  pdf_document:
    toc: yes
    toc_depth: '3'
  html_document:
    toc: yes
    toc_collapsed: yes
    toc_float: yes
    toc_depth: 3
    code_folding: show
    number_sections: no
    theme: lumen
    highlight: tango
---

# Libraries & data

```{r setup, echo=FALSE, show_col_types = FALSE}

library(tidyverse)

pat <- read_csv("D:\\PCT_EPO_counts\\patent_counts_direct_epo_pct.csv", show_col_types = FALSE)

```

# Tidy data


## Add 0 values for missing years

Here I assume that years that are not present = years with no patents filed under PCT 

```{r,eval=FALSE, echo=FALSE}

years = (1977:2020)

df2 <- data.frame(matrix(ncol = 3, nrow = 0)) # create data frame with 0 rows and 3 columns

for (state in unique(pat$Up_reg_label)){ 
  for (i in (1:length(years))){
    df2 <- rbind(df2, c(state, years[i], 0))  # create empty df with state and years
  }
}
colnames(df2) <- c('Up_reg_label','prio_year','inv_share') # provide column names
for(i in (1:nrow(pat))){
  for (j in (1:nrow(df2))){
    if(pat[i,1] == df2[j,1] && pat[i,2] == df2[j,2]){ # if year is present, add patent count value to the panel
      df2[j,3] = pat[i,3] 
    }
  }
}

df2$inv_share <- as.numeric(df2$inv_share)  # transform to numeric

pat <- df2 # change to original n


pat <- pat %>% 
  rename(region = Up_reg_label) %>% 
  rename(patent_count = inv_share) 

pat$prio_year <- as.integer(pat$prio_year)

```

## Montana: change name

```{r}
for (i in 1:nrow(pat)){
  if (pat[i, 1] == "Montana (US)"){
    pat[i,1] <- "Montana"}
}
```

## Remove not regionalised patent counts 
```{r}
pat %>% 
  filter(region == "United States - not regionalised") %>% 
  summarise(across(patent_count, .fns = list(sum = sum)))

pat <- pat %>% 
  filter(region != "United States - not regionalised") 
```


# Add controls

## GDP - 1999-2020

**Real GDP (millions of chained 2012 dollars)**

Unit: Millions of chained 2012 dollars

Source: BEA (Bureau of Economic Analysis)
```{r,eval=FALSE, echo=FALSE}

gdp <- read_csv("D:/Controls/SAGDP1__ALL_AREAS_1997_2021.csv", show_col_types = FALSE)

gdp_filt <- gdp %>% 
  select(!c("GeoFIPS", "Region", "TableName", "LineCode", "IndustryClassification")) %>% 
  filter(Description %in% c("Real GDP (millions of chained 2012 dollars)")) %>% # "Current-dollar GDP (millions of current dollars)"
  filter(!GeoName %in% c("United States")) %>% 
  pivot_longer(cols = c("1997", "1998", "1999","2000" ,"2001" ,"2002" ,"2003" ,"2004" ,"2005" ,"2006" ,"2007" ,"2008" ,"2009" ,"2010" ,"2011" , "2012","2013","2014","2015","2016","2017","2018","2019","2020", "2021"), 
               names_to = "year", 
               values_to = "gdp") %>% 
  select(!c("Description", "Unit")) 

gdp_filt$year <- as.numeric(gdp_filt$year)

pat_gdp <- left_join(pat, gdp_filt, by = c("region" = "GeoName", "prio_year" = "year"))

# write_csv(pat_gdp,"D:\\pat_gdp.csv")

```

## Population - 1999-2020

**Resident population including Armed Forces**

Unit of Measure: thousand

Sources: Table by the U.S. Energy Information Administration, based on population data from the U.S. Census Bureau

```{r,eval=FALSE, echo=FALSE}
library(readxl)
pop <- read_excel("D:\\Controls\\use_oi.xlsx", sheet = 2, skip = 2)

us_codes <- as.data.frame(state.abb)

us_codes$name <- state.name

pat_codes <- left_join(pat_gdp, us_codes, by = c("region" = "name"))

for (i in 1:nrow(pat_codes)){             # add District of Columbia code
  if (pat_codes[i, 1] == "District of Columbia"){
    pat_codes[i,6] <- "DC"}
}

pop_filt <-  pop[,c(1,39:62)]

pop_long <- pop_filt %>% pivot_longer(cols = c("1997", "1998", "1999","2000" ,"2001" ,"2002" ,"2003" ,"2004" ,"2005" ,"2006" ,"2007" ,"2008" ,"2009" ,"2010" ,"2011" , "2012","2013","2014","2015","2016","2017","2018","2019","2020"), 
               names_to = "year", 
               values_to = "pop")

pop_long$year <- as.numeric(pop_long$year)

pat_gdp_pop <- left_join(pat_codes, pop_long, by = c("state.abb" = "State", "prio_year" = "year"))

pat_gdp_pop <- pat_gdp_pop %>% 
  rename("state_name"= "region") %>% 
  rename("state_code" = "state.abb") %>% 
  select(c("state_name", "state_code", "prio_year", "patent_count", "gdp", "pop"))

# write_csv(pat_gdp_pop,"D:\\pat_gdp_pop.csv")

```


## Crude oil prices 1999-2020

**Domestic Crude Oil First Purchase Prices by Area**

Unit: Dollars per Barre

Definitions:

* Domestic: Crude oil produced in the U.S. or from its "outer continental shelf" as defined in 43 U.S.C. 1331.
* Crude Oil: A mixture of hydrocarbons that exists in liquid phase in natural underground reservoirs and remains liquid at atmospheric pressure after passing through surface separating facilities. Liquids produced at natural gas processing plants are excluded. Crude oil is refined to produce a wide array of petroleum products, including heating oils; gasoline, diesel and jet fuels; lubricants; asphalt; ethane, propane, and butane; and many other products used for their energy or chemical content.
* First purchase price: An equity (not custody) transaction involving an arms-length transfer of ownership of crude oil associated with the physical removal of the crude oil from a property (lease) for the first time. A first purchase normally occurs at the time and place of ownership transfer where the crude oil volume sold is measured and recorded on a run ticket or other similar physical evidence of purchase. The reported cost is the actual amount paid by the purchaser, allowing for any adjustments (deductions or premiums) passed on to the producer or royalty owner.

Numbers are aggregated by Petroleum Administration for Defense District (PADD):

* PADD 1 (East Coast):
   - PADD 1A (New England): Connecticut, Maine, Massachusetts, New Hampshire, Rhode Island, Vermont.
   - PADD 1B (Central Atlantic): Delaware, District of Columbia, Maryland, New Jersey, New York, Pennsylvania.
   - PADD 1C (Lower Atlantic): Florida, Georgia, North Carolina, South Carolina, Virginia, West Virginia.
* PADD 2 (Midwest): Illinois, Indiana, Iowa, Kansas, Kentucky, Michigan, Minnesota, Missouri, Nebraska, North Dakota, Ohio, Oklahoma, South Dakota, Tennessee, Wisconsin.
* PADD 3 (Gulf Coast): Alabama, Arkansas, Louisiana, Mississippi, New Mexico, Texas, Federal Offshore Gulf.
* PADD 4 (Rocky Mountain): Colorado, Idaho, Montana, Utah, Wyoming.
* PADD 5 (West Coast): Alaska (North Slope and Other Mainland), Arizona, California, Hawaii, Nevada, Oregon, Washington, Federal Offshore California.

Source: U.S. Energy Information Administration, ([link to page](https://www.eia.gov/dnav/pet/pet_pri_dfp1_k_m.htm))

```{r, eval=FALSE, echo=FALSE}

state_names <- c("Alabama", "Alaska", "Arizona", "Arkansas","California","Colorado","Connecticut", "Delaware", "Florida","Georgia","Idaho","Illinois","Indiana","Iowa","Kansas","Kentucky","Louisiana","Maine","Maryland","Massachusetts","Michigan","Minnesota","Mississippi","Missouri","Montana","Nebraska","Nevada","New Hampshire","New Jersey", "New York", "New Mexico", "North Carolina", "North Dakota","Ohio", "Oklahoma", "Oregon", "Pennsylvania", "Rhode Island", "South Carolina","South Dakota", "Tennessee", "Texas", "Utah", "Vermont", "Virginia", "Washington","West Virginia","Wisconsin","Wyoming", "District of Columbia", "Hawaii")

library(readxl)
fuel <- read_excel("D:\\Controls\\fuel_prices\\PET_PRI_DFP1_K_A.xls", sheet = 2, skip = 2)

fuel_filt <- fuel %>% 
  select(c("Date", "U.S. Crude Oil First Purchase Price (Dollars per Barrel)", "East Coast (PADD 1) Crude Oil First Purchase Price (Dollars per Barrel)", "Midwest (PADD 2) Crude Oil First Purchase Price (Dollars per Barrel)", "Gulf Coast (PADD 3) Crude Oil First Purchase Price (Dollars per Barrel)", "Rocky Mountain (PADD 4) Crude Oil First Purchase Price (Dollars per Barrel)", "West Coast (PADD 5) Crude Oil First Purchase Price (Dollars per Barrel)"))

PADD1 = c('Connecticut','Maine','Massachusetts', 'New Hampshire', 'Rhode Island', 'Vermont', 'Delaware', 'District of Columbia', 'Maryland', 'New Jersey', 'New York', 'Pennsylvania','Florida', 'Georgia', 'North Carolina', 'South Carolina', 'Virginia', 'West Virginia')
PADD2 = c('Illinois', 'Indiana', 'Iowa', 'Kansas', 'Kentucky', 'Michigan', 'Minnesota', 'Missouri', 'Nebraska', 'North Dakota', 'Ohio', 'Oklahoma', 'South Dakota', 'Tennessee', 'Wisconsin')
PADD3 = c('Alabama', 'Arkansas', 'Louisiana', 'Mississippi', 'New Mexico', 'Texas')
PADD4 = c('Colorado', 'Idaho', 'Montana', 'Utah', 'Wyoming')
PADD5  = c('Alaska', 'Arizona', 'California', 'Hawaii', 'Nevada', 'Oregon', 'Washington')

fuel2 = fuel_filt[(100:122),]
years = as.numeric(1999:2021)
fuel2 = as.data.frame(fuel2)
df <- data.frame(matrix(ncol = 3, nrow = 0))

for (state in state_names){
  for (i in (1:length(years))){
    df <- rbind(df, c(state, years[i], 0))
  }
}
colnames(df) <- c('State','Year','fuel')

values = c()

for (state in state_names){
  if(state %in% PADD1){
    values = append(values, fuel2[,3])
  }
  if(state %in% PADD2){
    values = append(values, fuel2[,4])
  }
  if(state %in% PADD3){
    values = append(values, fuel2[,5])
  }
  if(state %in% PADD4){
    values = append(values, fuel2[,6])
  }
  if(state %in% PADD5){
    values = append(values, fuel2[,7])
  }
}


# Check if all states are there

PADD = c('Connecticut','Maine','Massachusetts', 'New Hampshire', 'Rhode Island', 'Vermont',
'Delaware', 'District of Columbia', 'Maryland', 'New Jersey', 'New York', 'Pennsylvania',
'Florida', 'Georgia', 'North Carolina', 'South Carolina', 'Virginia', 'West Virginia',
'Illinois', 'Indiana', 'Iowa', 'Kansas', 'Kentucky', 'Michigan', 'Minnesota', 'Missouri', 'Nebraska', 'North Dakota', 'Ohio', 'Oklahoma', 'South Dakota', 'Tennessee', 'Wisconsin',
'Alabama', 'Arkansas', 'Louisiana', 'Mississippi', 'New Mexico', 'Texas', 'Colorado', 'Idaho', 'Montana', 'Utah', 'Wyoming', 'Alaska', 'Arizona', 'California', 'Hawaii', 'Nevada', 'Oregon', 'Washington')


for(state in state_names){
  if (!(state %in% PADD)){
    print(state)
  }
}

df$fuel <- values # cbind values to df

df$Year <- as.numeric(df$Year)

pat_gdp_pop_fuel <- left_join(pat_gdp_pop, df, by = c("state_name" = "State", "prio_year" = "Year")) # merge data 

# write_csv(pat_gdp_pop_fuel,"D:\\pat_gdp_pop_fuel.csv")

```

  
## Average temperature 

Unit: Degrees Farheneit

Source: NOAA National Centers for Environmental information, Climate at a Glance: Global Time Series, published February 2023, retrieved on March 3, 2023 from https://www.ncei.noaa.gov/access/monitoring/climate-at-a-glance/global/time-series

For Hawaii:

Unit: Monthly Mean Avg Temperature for Honolulu Area, HI

Source: https://www.weather.gov/wrh/Climate?wfo=hfo
```{r, echo=FALSE,eval=FALSE}

library(janitor)

state_name <- c("alabama", "alaska", "arizona", "arkansas","california","colorado","connecticut","delaware","florida","georgia", "idaho","illinois","indiana","iowa","kansas","kentucky","louisiana","maine","maryland","massachusetts","michigan","minnesota","mississippi","missouri","montana","nebraska","nevada","new_hampshire","new_Jersey","new_mexico", "new_york", "north_carolina","north_dakota","ohio", "oklahoma","oregon","pennsylvania","rhode_island","south_carolina",
"south_dakota", "tennessee","texas","utah","vermont","virginia","washington","west_virginia","wisconsin","wyoming", "district_of_columbia")

csv <- c()
df1 <- data.frame(matrix(ncol = 3, nrow = 0))
avg_temp <- data.frame()

for (i in state_name){
  csv <- read_csv(paste0('D:\\Controls\\avg_temp\\1999-2021_',i,'.csv')) # read csv w i = name of state
  df <- csv[-c(1:3),1:2]  # select specific rows and cols
  
  df1 <- df %>% 
    row_to_names(row_number = 1) # 1 row to colname 
  
  df1$Date <-substr(df1$Date,1,nchar(df1$Date)-2) # remove last 2 digits of year var
  
  df1$State <- i # add column w state name 
  
  avg_temp <- rbind(avg_temp, df1) # append dataframe
}

avg_temp$State <- avg_temp$State %>% 
  str_replace_all("_", " ") %>%  # change names with _ with " 
  str_to_title() # put first case high

for (i in 1:nrow(avg_temp)){ # correct district of columbia  
  if (avg_temp[i, 3] == "District Of Columbia"){
    avg_temp[i, 3] <- "District of Columbia"
  }
}

# hawaii

hawaii <- read_csv('D:\\Controls\\honolulu_area_avg_temp.csv')

hawaii <- hawaii %>% 
  mutate(State = "Hawaii") %>% 
  rename("Date" = "Year") %>% 
  rename("Value" = "Annual")

avg_temp_all <- rbind(avg_temp, hawaii)

avg_temp_all$Date <- as.integer(avg_temp_all$Date)

# join with all data

pat_gdp_pop_fuel_temp <- left_join(pat_gdp_pop_fuel, avg_temp_all, by = c("state_name" = "State", "prio_year" = "Date"))

df <- pat_gdp_pop_fuel_temp %>% 
  rename("avg_temp" = "Value")
```

# Sanity check - complete dataset 

```{r}

df %>% filter(!complete.cases(.)) 

write_csv(df,"D:\\For_model\\final_df.csv")

```

