---
title: "Model specification pipeline"
author: "Laura Menicacci"
date: "`r Sys.Date()`"
output: html_document
---

# Libraries

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(gets)
library(getspanel)
library(here)
library(doParallel)
library(readxl)
library(gdata)

```

# Create controls

```{r}

df <- read_csv(".\\df.csv", show_col_types = FALSE)

df2 <- df %>% 
  mutate(lpop = log(pop), # log pop
         lgdp = log(gdp), # log gdp
         lfuel = log(fuel), # log fuel price by purchase area 
         lgdp_sq = log(gdp)^2, # squared gdp
         ltemp = log(avg_temp), # average temperature
         ihs_patent_count = asinh(patent_count), # inverse hyperbolic sine transformation
         log_patent_count = log(patent_count+1)) %>% # log+1 transformation
  filter(prio_year >= 2000)

df2 %>% pull(state_name) %>% unique # sanity check

df2 %>% filter(!complete.cases(.))

# write_csv(df2, ".\\df.csv")

```

# Load samples

Different samples are tried out that remove states according to quantile distributions. q30 is the one chosen: removes states up to 30th quantile of distribution. 
```{r}
us_all <- c("Alabama", "Alaska", "Arizona", "Arkansas","California","Colorado","Connecticut", "Delaware", "Florida","Georgia","Idaho","Illinois","Indiana","Iowa","Kansas","Kentucky","Louisiana","Maine","Maryland","Massachusetts","Michigan","Minnesota","Mississippi","Missouri","Montana","Nebraska","Nevada","New Hampshire","New Jersey", "New York", "New Mexico", "North Carolina", "North Dakota","Ohio", "Oklahoma", "Oregon", "Pennsylvania", "Rhode Island", "South Carolina","South Dakota", "Tennessee", "Texas", "Utah", "Vermont", "Virginia", "Washington","West Virginia","Wisconsin","Wyoming", "District of Columbia", "Hawaii")

us_q20 <- c("Alabama", "Arizona", "Arkansas","California","Colorado","Connecticut", "Delaware", "Florida","Georgia","Idaho","Illinois","Indiana","Iowa","Kansas","Kentucky","Louisiana","Maine","Maryland","Massachusetts","Michigan","Minnesota","Missouri","Nebraska","Nevada","New Hampshire","New Jersey", "New York", "New Mexico", "North Carolina", "Ohio", "Oklahoma", "Oregon", "Pennsylvania", "Rhode Island", "South Carolina","South Dakota", "Tennessee", "Texas", "Utah", "Vermont", "Virginia", "Washington","Wisconsin", "District of Columbia", "Hawaii")

us_q25 <- c("Alabama", "Arizona", "California","Colorado","Connecticut", "Delaware", "Florida","Georgia","Idaho","Illinois","Indiana","Iowa","Kansas","Kentucky","Louisiana","Maine","Maryland","Massachusetts","Michigan","Minnesota","Missouri","Nevada","New Hampshire","New Jersey", "New York", "New Mexico", "North Carolina", "Ohio", "Oklahoma", "Oregon", "Pennsylvania", "Rhode Island", "South Carolina","South Dakota", "Tennessee", "Texas", "Utah", "Vermont", "Virginia", "Washington","Wisconsin", "District of Columbia", "Hawaii")

us_q30 <- c( "Arizona","California","Colorado","Connecticut", "Delaware", "Florida","Georgia","Idaho","Illinois","Indiana","Iowa","Kentucky","Maryland","Massachusetts","Michigan","Minnesota","Missouri","Nevada","New Hampshire","New Jersey", "New York", "New Mexico", "North Carolina", "Ohio", "Oklahoma", "Oregon", "Pennsylvania", "Rhode Island", "South Carolina", "Tennessee", "Texas", "Utah","Virginia", "Washington","Wisconsin", "District of Columbia")

samples <- mget(c("us_all", "us_q20", "us_q25", "us_q30"))
```

# Functional forms

```{r}
controls <- c(" ~ lgdp + lpop + lgdp_sq ") # specify controls to be fed into model 

dep_var <- c("log_patent_count") # specify dependent var to be fed into model 

base_forms <- paste0(rep(dep_var, each = length(controls)), controls) # paste together

base_forms

```

# Analysis: main model parameters

```{r}

cl <- makeCluster(6) 
registerDoParallel(cl) # foreach package function that parallelizes operations to make it faster using %:% 

models <- foreach(f = base_forms, .combine = rbind, .packages = c('tidyverse', 'getspanel')) %:% # specify functional form 
  foreach(smpl = c("us_q30"), .combine = rbind) %:% # specify samples
  foreach(a = c(0,1), .combine = rbind) %:%
  foreach(p.value = c(0.01, 0.05), .combine = rbind, .errorhandling = "remove") %dopar% {
    dat <- df2 %>% filter(state_name %in% samples[[smpl]]) %>% filter(prio_year <= 2019) # iterate over samples 
      is <- isatpanel( # main function
            data = dat,
            formula = as.formula(f),
            index = c("state_name", "prio_year"), # id and time
            effect = "twoways", # twoway fixed effects chosen as estimator
            iis = TRUE, # enable impulse indicator saturation
            fesis = TRUE, # enable fixed effects indicator saturation
            ar = a, # specify auto-regressive terms
            t.pval = p.value,  # false positive rate
            max.block.size = 20 # size for block search 
            )
          models = tibble(source = f,  # results are gathered into a dataframe with the following information:
                          state_sample = smpl,    # which sample 
                          year_range = paste0(min(dat$prio_year),":",max(dat$prio_year)), # time series
                          p_val = p.value, # false positive rates 
                          is = list(is), # here there's the whole model output with results of interest
                          iis = TRUE, # whether IIS was enabled or not
                          b_size = 20,
                          ar = a)
  }

stopCluster(cl) # stop parallelizing 

saveRDS(models, ".\\results\\mods_main_updatedpackage.RDS")  # save model output 
```

## Analysis: silenced IIS 

```{r}
cl <- makeCluster(6)
registerDoParallel(cl)

models <- foreach(f = base_forms, .combine = rbind, .packages = c('tidyverse', 'getspanel')) %:%
  foreach(smpl = c("us_q30"), .combine = rbind) %:%
  foreach(a = c(0,1), .combine = rbind) %:%
  #foreach(b = c(10, 20, 30), .combine = rbind) %:%
  foreach(p.value = c(0.01, 0.05), .combine = rbind, .errorhandling = "remove") %dopar% {
    dat <- df2 %>% filter(state_name %in% samples[[smpl]]) %>% filter(prio_year <= 2019)
      is <- isatpanel(
            data = dat,
            formula = as.formula(f),
            index = c("state_name", "prio_year"),
            effect = "twoways",
            iis = FALSE,
            fesis = TRUE,
            ar = a,
            t.pval = p.value,
            max.block.size = 20
            )
          models = tibble(source = f, 
                          state_sample = smpl,      
                          year_range = paste0(min(dat$prio_year),":",max(dat$prio_year)),
                          p_val = p.value, 
                          is = list(is),
                          iis = FALSE,
                          b_size = 20,
                          ar = a)
  }

stopCluster(cl)

saveRDS(models, ".\\results\\mods_main_updatedpackage_noIIS.RDS")
```


## Analysis: try out different robustness checks

Selected model specifications + includes robustness checks controls: fuel

```{r}

# Functional forms

add_forms <- paste0(base_forms, "+ lfuel")

# Initialize model

cl <- makeCluster(6)
registerDoParallel(cl)

models <- foreach(f = add_forms, .combine = rbind, .packages = c('tidyverse', 'getspanel')) %:%
  foreach(smpl = c("us_all", "us_q30"), .combine = rbind) %:%
  foreach(a = c(0,1), .combine = rbind) %:%
  #foreach(b = c(10, 20, 30), .combine = rbind) %:%
  foreach(p.value = c(0.01, 0.05), .combine = rbind, .errorhandling = "remove") %dopar% {
    dat <- df2 %>% filter(state_name %in% samples[[smpl]]) %>% filter(prio_year <= 2019)
      is <- isatpanel(
            data = dat,
            formula = as.formula(f),
            index = c("state_name", "prio_year"),
            effect = "twoways",
            iis = TRUE,
            fesis = TRUE,
            ar = a,
            t.pval = p.value,
            max.block.size = 20
            )
          models = tibble(source = f, 
                          state_sample = smpl, 
                          year_range = paste0(min(dat$prio_year),":",max(dat$prio_year)), 
                          p_val = p.value, 
                          is = list(is),
                          iis = TRUE,
                          b_size = 20,
                          ar = a)
  }


saveRDS(models, ".\\results\\models_w_fuel.RDS")

```

