---
title: "thesis_data_analysis"
author: "Laura Menicacci"
date: "2023-02-01"
output: html_document
---

# libraries
```{r setup, include=FALSE}
library(tidyverse)
```

# Data

## Patent by technology

from https://stats.oecd.org/index.aspx?queryid=29068# 

data selected:
- technology domain: CLIMATE CHANGE MITIGATION technologies related to ENERGY GENERATION, TRANSMISSION OR DISTRIBUTION (IPC class Y02E) --> cannot select specific tech 
- country sample: all OECD + BRIICS economies
- selected both priority (date when the first patent for a specific technology was granted) and application date
- Value: number of patents using country's fractional value
- year span: 1976-2020

```{r}
pat <- read_csv("D:\\Aggregated\\PATS_IPC_01022023161250362.csv") 

# pat$Country %>% unique()

EU_15 <- c("AUT","BEL", "DNK", "FIN", "FRA", "DEU", "GRC",
                      "IRL", "ITA", "LUX", "NLD", "PRT", "ESP", "SWE", "GBR")

top_3 <- c("JPN", "USA", "DEU")

pat %>% 
  filter(Country %in% c("Germany")) %>%  # example with Germany
    filter(`Technology domains & IPC` %in% c("Climate change mitigation technologies related to energy generation, transmission or distribution")) %>% 
  ggplot(aes(x = Time, y = Value, group = `Reference Date`, fill = `Reference Date`)) + geom_col(position = "dodge", width = 0.8) + theme_bw() + scale_fill_brewer(palette = "Paired") + ggtitle("Number of Y02E German patents by Reference Date")

pat %>% 
  filter(LOCATION %in% EU_15) %>% 
  filter(`Reference Date` %in% c("Priority date")) %>%
  filter(`Patents Office & Patents Families` %in% c("Patent applications to the EPO")) %>% 
  filter(IPC %in% c("TOTAL")) %>% 
  ggplot(aes(x = Time, y = Value, colour = LOCATION)) + geom_line() + theme_bw() + ggtitle("Total patents over time in EU15")

pat %>% 
  filter(LOCATION %in% EU_15) %>% 
  filter(`Reference Date` %in% c("Priority date")) %>%
  filter(`Patents Office & Patents Families` %in% c("Patent applications to the EPO")) %>% 
  filter(`Technology domains & IPC` %in% c("Climate change mitigation technologies related to energy generation, transmission or distribution")) %>% 
  ggplot(aes(x = Time, y = Value, colour = LOCATION)) + geom_line() + theme_bw() + ggtitle("Y02E patents over time in EU15")

pat %>% 
  filter(LOCATION %in% top_3) %>% 
  filter(`Reference Date` %in% c("Priority date")) %>%
  filter(`Patents Office & Patents Families` %in% c("Patent applications to the EPO")) %>% 
  filter(IPC %in% c("TOTAL")) %>% 
  ggplot(aes(x = Time, y = Value, colour = LOCATION)) + geom_line() + theme_bw() + ggtitle("Total patents over time in top 3 innovators")

pat %>% 
  filter(LOCATION %in% top_3) %>% 
  filter(`Reference Date` %in% c("Priority date")) %>%
  filter(`Patents Office & Patents Families` %in% c("Patent applications to the EPO")) %>% 
  filter(`Technology domains & IPC` %in% c("Climate change mitigation technologies related to energy generation, transmission or distribution")) %>% 
  ggplot(aes(x = Time, y = Value, colour = LOCATION)) + geom_line() + theme_bw() + ggtitle("Y02E patents over time in top 3 innovators")

```
## Technology development

from https://stats.oecd.org/Index.aspx?DataSetCode=PAT_DEV#

data selected:
- technology: CLIMATE CHANGE MITIGATION technologies related to ENERGY GENERATION, TRANSMISSION OR DISTRIBUTION (IPC class Y02E) --> possible to select specific tech for ex. wind energy, solar energy, etc
- family size: 1, 2, 3, 4 and greater --> 4 and greater are more important patents
- type of variable: number of patent with country fractional value
- country sample: OECD + BRIICS
- year span: 1960-2019

```{r}
pat_dev <- read_csv("D:\\Aggregated\\PAT_DEV_02022023120219870.csv")


pat_dev %>% 
  filter(COU %in% EU_15) %>% # example with EU 15
  filter(`Family size` %in% c("Four and greater")) %>% 
  ggplot(aes(x = Year, y = Value, colour = COU)) + geom_line() + theme_bw() + 
  ggtitle("Y02E patents for EU 15 - family size: 4 and greater")

pat_dev %>% 
  filter(COU %in% top_3) %>% # example with top 3 innovators
  filter(`Family size` %in% c("Four and greater")) %>% 
  ggplot(aes(x = Year, y = Value, colour = COU)) + geom_line() + theme_bw() + ggtitle("Y02E patents for the top 3 innovators - family size: 4 and greater")


```

## Citation and Quality data from OECD STI Microdata Lab

- from https://www.oecd.org/sti/intellectual-property-statistics-and-analysis.html
- only opening EPO data as an example but there's more

```{r}
##qual <- read.csv("D:\\OECD_PATENT_QUALITY_EPO_INDIC_3.csv", sep = "|") # best option: tried fwf, _csv, #tried opening without python prepro
#
#q_sample <- qual[1:3000,]
#
#q_sample %>% names()
#
##cit_counts <- read.csv("D:\\EPO_CIT_COUNTS.csv", sep = "|")  # index still remains :(
#
#c_sample <- cit_counts[1:3000,]
#
#c_sample %>% names()

```
## Renewable energies

```{r}
ren <- read_csv("D:\\PAT_DEV_RE.csv")

tech <- c("Wind energy", "Solar thermal energy" , "Solar photovoltaic (PV) energy", "Solar thermal-PV hybrids", "Geothermal energy", "Marine energy, e.g. using wave energy or salinity gradient", "Hydro energy"  )

ren %>% 
  filter(COU %in% c("DEU")) %>% # example with top 3 innovators
  filter(`Family size` %in% c("Four and greater")) %>% 
  filter(`Technology domain` %in% tech) %>% 
  ggplot(aes(x = Year, y = Value, colour = `Technology domain`)) + geom_line() + theme_bw() + ggtitle("Y02E10 patents for Germany - family size: 4 and greater")


ren %>% 
  filter(COU %in% EU_15) %>% # example with top 3 innovators
  filter(`Family size` %in% c("Four and greater")) %>% 
  filter(`Technology domain` %in% c("Renewable energy generation")) %>% 
  ggplot(aes(x = Year, y = Value, colour = COU)) + geom_line() + theme_bw() + ggtitle("Y02E10 patents for EU10 - family size: 4 and greater")

ren %>% 
  filter(COU %in% top_3) %>% # example with top 3 innovators
  filter(`Family size` %in% c("Four and greater")) %>% 
  filter(`Technology domain` %in% c("Renewable energy generation")) %>% 
  ggplot(aes(x = Year, y = Value, colour = COU)) + geom_line() + theme_bw() + ggtitle("Y02E10 patents for EU10 - family size: 4 and greater")

```


```{r}

reg_oecd <- read_csv("D:\\Aggregated\\PATS_REGION_05022023190614585.csv")

reg_oecd %>%  # computing average of patent count by region
    filter(`Total patents and by technologies` %in% c("Climate change mitigation technologies related to energy generation, transmission or distribution")) %>% 
  group_by(Regions) %>% 
  summarise(avg = mean(Value)) %>% 
  arrange(desc(avg))

top_3 = c("San Jose-San Francisco-Oakland - CA", "Boston-Worcester-Manchester - MA-NH", "New York-Newark-Bridgeport - NY-NJ-CT-PA")

top_10 <- c("San Jose-San Francisco-Oakland - CA", "Boston-Worcester-Manchester - MA-NH", "New York-Newark-Bridgeport - NY-NJ-CT-PA", "Los Angeles-Long Beach-Riverside - CA", "Philadelphia-Camden-Vineland - PA-NJ-DE-MD", "Detroit-Warren-Flint - MI", "Albany-Schenectady-Amsterdam - NY","Hartford-West Hartford-Willimantic - CT", "Minneapolis-St. Paul-St. Cloud - MN-WI","Houston-Baytown-Huntsville - TX")

last_10 <- c("Pendleton-Hermiston - OR", "San Angelo - TX", "Bismarck - ND", "Redding - CA", "Monroe-Bastrop - LA", "Kearney - NE", "Aberdeen - SD", "Cape Girardeau-Jackson - MO-IL", "Paducah - KY-IL", "Minot - ND")

fig_top_10 <- reg_oecd %>% 
  #filter(`Patent office` %in% c("Patent applications to the EPO")) %>% 
  filter(`Reference region` %in% c("Inventor(s)'s country(ies) of residence")) %>% 
  filter(`Total patents and by technologies` %in% c("Climate change mitigation technologies related to energy generation, transmission or distribution")) %>% 
  filter(Regions %in% top_10) %>% 
  #group_by(Regions) %>% 
  ggplot(aes(x = Time, y = Value, colour = `Patent office`)) + geom_line(linewidth = 1) + theme_bw() + ggtitle("Y02E EPO vs PCT counts in top 10 patent-intensive US regions") + scale_x_discrete(limits = c(1998, 2000, 2002, 2004, 2006, 2008, 2010, 2012)) + facet_wrap(~Regions, scale="free_y") + theme(legend.position = "top")

fig_top_10

fig_top <- reg_oecd %>% 
  filter(`Patent office` %in% c("Patent applications to the EPO")) %>% 
  filter(`Reference region` %in% c("Inventor(s)'s country(ies) of residence")) %>% 
  filter(`Total patents and by technologies` %in% c("Climate change mitigation technologies related to energy generation, transmission or distribution")) %>% 
  filter(Regions %in% top_3) %>% 
  #group_by(Regions) %>% 
  ggplot(aes(x = Time, y = Value, colour = Regions)) + geom_line(linewidth = 1) + theme_bw() + ggtitle("Y02E EPO patents over time in top 3 patent-intensive regions") + scale_x_discrete(limits = c(1998, 2000, 2002, 2004, 2006, 2008, 2010, 2012))


fig_last_10 <- reg_oecd %>% 
  #filter(`Patent office` %in% c("Patent applications to the EPO")) %>% 
  filter(`Reference region` %in% c("Inventor(s)'s country(ies) of residence")) %>% 
  filter(`Total patents and by technologies` %in% c("Climate change mitigation technologies related to energy generation, transmission or distribution")) %>% 
  filter(Regions %in% last_10) %>% 
  #group_by(Regions) %>% 
  ggplot(aes(x = Time, y = Value, colour = `Patent office`)) + geom_line(linewidth = 1) + theme_bw() + ggtitle("Y02E EPO vs PCT counts in last 10 patent-intensive US regions") + scale_x_discrete(limits = c(1998, 2000, 2002, 2004, 2006, 2008, 2010, 2012)) + facet_wrap(~Regions, scale="free_y") + theme(legend.position = "top")


fig <- reg_oecd %>% 
  #filter(`Patent office` %in% c("Patent applications to the EPO")) %>% 
  filter(`Reference region` %in% c("Inventor(s)'s country(ies) of residence")) %>% 
  filter(`Total patents and by technologies` %in% c("Climate change mitigation technologies related to energy generation, transmission or distribution")) %>% 
  filter(Regions %in% c("United States")) %>% 
  #group_by(Regions) %>% 
  ggplot(aes(x = Time, y = Value, colour =`Patent office`)) + geom_line(linewidth = 1) + theme_bw() + ggtitle("Y02E patents over time in the US") + labs(subtitle = "By Patent Office") + scale_x_discrete(limits = c(1998, 1999, 2000, 2001, 2002, 2003, 2004, 2005, 2006, 2007, 2008, 2009, 2010, 2011, 2012, 2013))

fig 

fig_last_10
```

```{r}
nas <- which((reg_oecd$Value == "0"))

(table(reg_oecd$Regions[nas])) # Lewiston - ID-WA

table(reg_oecd$Time[nas])



```



```{r}
ggsave("Y02E_patents_byPO.png", plot = fig, device = "png", width = 10, height = 8)

ggsave("Y02E_patents_top.png", plot = fig_top, device = "png", width = 10, height = 8)

ggsave("Y02E_patents_last_10.png", plot = fig_last_10, device = "png", width = 10, height = 8)

ggsave("Y02E_patents_top10.png", plot = fig_top_10, device = "png", width = 12, height = 10)


```

